---
title: 'CLASS_06: R Functions'
author: "Berenice R. Leal"
date: "1/23/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions

## Level 2 heding

### Level 3 heading

```{r}
# This is a silly plot
plot(1:5)
```

Lets see more about **file import** (i.e. reading files into R). Th main read function in base R is 'read.table()'


```{r}
read.table("test1.txt")
```

```{r}
t1 <- read.table("test1.txt", sep=",", header=TRUE)
t1
```


Or I could just use 'read.csv()' which has the arguents I want in this case!



Second file
```{r}
read.table("test2.txt", sep="$", header=TRUE)
```

#Back to functions
Our first example function


```{r}
add <- function (x, y=1) { 
  # Sum the inout x and y
  x + y
}
```

Let's try using this function

```{r}
add(7,3)
```


How does this work with vectors as input
```{r}
add( c(1,2,4), 4)
```

```{r}
add( c(1,2,4), c(1,2,4))
```


What is this 'range()' function you talk of?
```{r}
x <- c(4,4,10,3,11)
max(x)
min(x)
```

```{r}
range(x)

```

This is our second function:

```{r}
rescale <- function(x) {
 rng <- range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}

rescale(x)

rescale (1:10)

#How would you get your function to work here...
rescale( c(1,2,NA,3,10))


```

```{r}
x <- c(1,2, NA, 3,10)
rng <- range(x)
rng
```
```{r}
rng <- range(x, na.rm = TRUE)
rng

```


```{r}
rescale2 <- function(x) {
 rng <- range(x, na.rm=TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}
```


```{r}
rescale2( c(1,2,NA,3,10) )
```


```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 
 print("dont sing again")
 
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```

```{r}
rescale3(x, plot=TRUE)
```



PART B

Hands on Section

```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```



```{r}
s1 <- read.pdb("4AKE") # kinase with drug
```

```{r}
s1

```
Q1. What type of object is returned from the 'read.pdb()' 
It is a list of 8 things and of class 

```{r}
class(s1)
str(s1)
```

Q2. What does the trim.pdb() function do?

```{r}
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1

```

Q4. What would be a better plot to compare across the different proteins? 
1 plot with all the graphs in it

```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, typ="l", col="blue", lwd=2)
points(s3.b, typ="l", col="red", lwd=2)



```

```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)

```


```{r}
library(bio3d)
```


```{r}

```


```{r}

s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")


```

```{r}

library(bio3d)
myplotfunction <- function(file, chain, elments, fctr) {
  
  # depicts data with different colors in the graph
  plot_colors <- c("blue", "pink", "green")
  
  
  # to iterate through values in file vector -read everything-
  for (x in 1:length(file)) {
  s1 <- read.pdb(file[x])

  # Just a small modification in one of the repetitive lines to make paste more clear
  s1.chain <- trim.pdb(s1, chain = chain, elety = elements)
  
  # allatoms takes in all the atom information and  
  # selects an entire column based on the factor input 
  atom_df <- s1.chain$atom

  allatoms <- atom_df[, fctr] 
  
  
  # creates the first plot
  if (x == 1) {
    plotb3(allatoms, sse = s1.chain, typ = "l", ylab = paste(toupper(fctr), "factor", sep = ""), col = plot_colors[x])
    
    # adds additional plots to first plot
  } else {
    lines(allatoms, col = plot_colors[x])
  }
  }
  
  # creates a legend for the graph
  legend("topleft",cex = 0.5, inset = c(0.05, 0.05), title = "Kinases - Drug Interaction", file, fill = plot_colors)
}

# definitions of what will be inside my function
files <- c("4AKE", "1AKE", "1E4Y")
chains <- "A"
elements <- "CA"
factors <- "b"

# Myfunction

myplotfunction(files, chains, elements, factors)



```



```{r}









```









